---
title: "hw5"
author: "jiayi lou"
date: "2021/2/9"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

#initial analysis

```{r}
library(car)
library(tidyverse)
str(carData::Vocab)#view the data type

library(ggplot2)# Beautiful drawing package

ggplot(Vocab, aes(x = factor(year), y = education,size = vocabulary )) +
  geom_point()

ggplot(Vocab, aes(x = factor(year), y = vocabulary,size = education)) + geom_point()

```

There are 4 variables in this data set. Our research focuses on the connection between words and other variables, and how to predict. It is difficult to see the pattern through above simple distribution plots.

## sex

```{r}
Vocab %>%
  ggplot(mapping = aes(x= sex,y=vocabulary ,fill=sex))+
  geom_violin()+
  geom_boxplot()
Vocab %>%
  ggplot(mapping = aes(x= sex,y=education ,fill=sex))+
  geom_violin()+
  geom_boxplot()
```
From the above figure, we can see that the average education level of men is higher than that of women. There is also a big difference in distribution. However,the average vocabulary level of men and women are almost the same. The distribution is also similar.

```{r}
Vocab %>%
  ggplot()+
  geom_bar(mapping = aes(x = factor(year), fill=sex),show.legend = TRUE,position="dodge")

Vocab %>%
  ggplot()+
  geom_bar(mapping = aes(x = education , fill=sex),show.legend = TRUE,position="dodge")

Vocab %>%
  ggplot()+
  geom_bar(mapping = aes(x= vocabulary , fill=sex,),show.legend = TRUE,position="dodge")
```


Through the histogram, we can see that there are more women than men.


## year

```{r}
Vocab %>%
  ggplot(aes(factor(year),vocabulary,fill=factor(year)))+
  geom_violin()+
  geom_boxplot()

Vocab %>%
  ggplot(mapping = aes(x= factor(year),y=education,fill=factor(year)))+
  geom_boxplot()+
  geom_violin()+
  geom_boxplot()

```

Excluding the consideration of gender, we found that as the years have grown, everyone’s education level has also risen slowly, but the vocabulary test score has remained at around 6.This makes people suspect that there is no connection between education and vocabulary test results.We will verify this suspicion in the deep analysis section.


```{r}
Vocab %>%
  ggplot(aes(x=education, fill=vocabulary)) +
  geom_density() +
  facet_grid(vocabulary~.)
```
Through the density map, we find that people with high education are more likely to get good vocabulary test scores than people with low education. So we can refute the previous view. Education have effects on vocabulary.

## education
```{r}

Vocab %>%
  ggplot()+
  geom_bar(mapping = aes(x = education, fill=factor(year)),show.legend = TRUE)
```


Through the histogram, we can see that there are more women than men, People with 12 levels of education account for a large proportion, and the number of illiterate samples is very small.



## vocabulary

```{r}
Vocab %>%
  ggplot(aes(x = vocabulary,fill=sex))+
  geom_bar(position="dodge")

```


# regression analysis

## simple regression models without interaction

```{r}
model1<-lm(vocabulary~year+education+sex,Vocab)
summary(model1)
```


```{r}
spida2::wald(model1)
```

For simple regression models without interaction, we don’t need to drop any variables.



## year
```{r}
model.year<-lm(vocabulary~year,Vocab)
summary(model.year)

ggplot(Vocab,aes(year, vocabulary,colour = sex)) +
  ggtitle("male vs female")+
  geom_point() +
  stat_ellipse(type = "norm", linetype = 1, level=c(0.68) )+
  stat_ellipse(type = "norm", linetype = 2, level=c(0.95) )+
  coord_fixed()+
  geom_smooth(method = "lm",linetype=1,se=FALSE,colour="red")

ggplot(Vocab,aes(year, vocabulary)) +
  ggtitle("mixed")+
  geom_point() +
  stat_ellipse(type = "norm", linetype = 1, level=c(0.68) )+
  stat_ellipse(type = "norm", linetype = 2, level=c(0.95) )+
  coord_fixed()+
  geom_smooth(method = "lm",linetype=1,se=FALSE,colour="red")
```

It seems there is no connection between year and vocabulary,and confidence ellipse is not titrated.


## education
```{r}
model.vecabulary<-lm(vocabulary~year*sex,Vocab)
summary(model.vecabulary)

ggplot(Vocab,aes(education,vocabulary,colour = sex)) +
  geom_point() +
  ggtitle("male vs female")+
  stat_ellipse(type = "norm", linetype = 1, level=c(0.68) )+
  stat_ellipse(type = "norm", linetype = 2, level=c(0.95) )+
  coord_fixed()+
  geom_smooth(method = "lm",linetype=1,se=FALSE,colour="red")

ggplot(Vocab,aes(education,vocabulary)) +
  geom_point() +
  ggtitle("mixed")+
  stat_ellipse(type = "norm", linetype = 1, level=c(0.68) )+
  stat_ellipse(type = "norm", linetype = 2, level=c(0.95) )+
  geom_smooth(method = "lm",linetype=1,se=FALSE,colour="red")


```


Education has positive effect on vocabulary,and confidence ellipse is titrated.



# model with interaction


```{r}
model.all<-lm(vocabulary~year*education*sex,Vocab)
summary(model.all)

# model with 3-way interaction


```{r}
model.all<-lm(vocabulary~year*education*sex,Vocab)
summary(model.all)

anova(model.all,lm(vocabulary~year*education+sex*education+year*sex,Vocab),test='LRT')#Three way interaction term shows significance. 

crPlots(lm(vocabulary~education+sex+year,data=Vocab)) #crPlots suggest that a squared term for education may be of interest
#An anova test comparing the model with all interaction terms against a model with all interactions plus the squared education term produces a small values, suggesting #significance. 
anova(model.all,lm(vocabulary~sex*education*year+I(education^2),Vocab),test='LRT') 

#Based on the box and whisker plots, women were initially MORE educated then men at a minimum (excluding outliers), whereas men were MORE educated than women at 
#a maximum. That is, the lowest educated women were more edcuated then the lowest educated men, and the highest educated women were less educated than the 
#highest educated men, excluding outliers. It wasn't until around 2004 that the education gap between men and women included in the vocabulary study
#evened out. For the most part, the average education between men and women is about equal; it is slighlty higher for men from 1989-1993. The average level of 
#edcuation for women is higher only in 2006. 
bwplot(education~sex|as.factor(year),Vocab,groups=sex)

#Based on the box and whisker plots comparing sex against vocabulary, men and women have, on average, equal vocabulary levels each year. In the years 1976, 
#1978, 1984, and 1988 the minimum vocabulary score was higher for women than men, excluding outliers. In the years 1974, 1994, and 1998 the minimum score was #higher for men #than women. The maximum scores are equal for men and women #throughout all years. 
bwplot(vocabulary~sex|as.factor(year),Vocab,groups=sex)

#There were more women then men particpiating in the survey in every single year that the vocabulary survey was performed
up(Vocab,~sex+year)


```
